Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-10-25 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

As a benchmark we start by considering a single sphere within an 
optical trap. A single polystyrene sphere suspended in water 
($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focus 
Gaussian beam and its trajectory was recorded. A 3 second is 
trajectory is a typical measurement time for collecting a power 
spectrum, the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{PSD_sphere.png}
	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian},
		 scattered points represents the blocked data ($n_b=100$). Corner 
		 frequency for the Lorentzian curves are reported in the legend.}
	\label{fig:psd_sphere}
\end{figure}

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra report 
different corner frequencies which would be indicate that the trap 
is not perfectly circular. Using \textit{ott} we can compare the 
expected trap strength to what is reported by a quadrant photo diode:
\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

Where the ellipticity of the beam is given by $e = (1-\kappa_y/
\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
wavefront. Its clear from these initial results that the QPD 
is more sensitive to changes along the y-axis than the x-axis 
when compared to the direct \textit{ott} calculations. Typically, 
even an industrial Gaussian beam will produce an elliptical 
diffraction limit spot when heavily focused; in their tutorial 
for optimizing the PSD analysis, Berg and Sorensen reported a 
ellipticity of around 15 \% after a total calibration time of 
81 seconds \cite{BergSoerensen2004}. The reason for this 
discrepancy can be explained partly by the fact that our 
estimation of the trap geometry via \textit{ott} is based on 
the force-displacement curve while the sphere is moving in only 
one direction, whereas the QPD is estimating the trap geometry 
by extrapolating from far field scattering signals. We now 
consider a symmetric dimer in identical simulative conditions.

\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
		\hline
		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
		& \parbox{1.25cm}{\centering 274} & 285 \\
		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{42.8 \%} &
		\multicolumn{2}{|c|}{12.7 \%} & 
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
Considering the simplicity of a scatterer such as a dimer,
one would assume that the dynamics of such an object would be
relatively easy to predict. Simulations of dimers in the Mie regime
show that not only do they have multiple positions and orientations 
in which they can be trapped but also that their interaction
with circularly polarised light is heavily dependent on the axial 
position and trapping orientation. Dimer's have the potential to be used
as adaptive micro-rotors, being simple to synthesise and can be made 
out of any material of choice, but in order for these particles to be
used as such one needs a means of characterising the optical torque
applied. While estimations can be made in the Rayleigh regime due to 
the dimer being treated as point dipole, Mie regime micro rotors require 
an in-depth theoretical description of the mechanism that is creating 
this optical torque.

>>>>

As a benchmark we start by considering a single sphere within an 
optical trap. A single polystyrene sphere suspended in water 
($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
Gaussian beam ($NA=1.25$) using circularly polarised light. For 
the sake of time efficiency the trajectory was sampled every 10
time steps, meaning the upper bound on the power spectra is
$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
we fitted the power spectra using the aliased Lorentzian 
(Eq.~\eqref{eq:aliased_lorentzian}).
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{PSD_sphere.png}
	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian},
		 scattered points represents the blocked data ($n_b=100$). Corner 
		 frequency for the Lorentzian curves are reported in the legend.}
	\label{fig:psd_sphere}
\end{figure}

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra report 
different corner frequencies which would be indicate that the trap 
is not perfectly circular. We can use both \textit{ott} and the 
trajectory itself to derive an estimation of the trap geometry. 
The corner frequencies and corresponding trap stiffness are 
reported below:
\newpage

\begin{center}
	\captionof{table}{QPD fitting for single sphere}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline

	\end{tabular}
\end{center}

Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, where
the variance is based upon our choice of frequency boundaries
({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And the ellipticity 
of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a 
measure of the symmetry of the beam wavefront. Its clear from these 
initial results that the QPD is more sensitive to changes along the 
y-axis than the x-axis when compared to the direct \textit{ott} 
calculations. This is somewhat reflected in the trajectory results. 
Typically, even an industrial Gaussian beam will produce an 
elliptical diffraction limit spot when heavily focused; in their 
tutorial for optimizing the PSD analysis, Berg and Sorensen reported 
a ellipticity of around 15 \% after a total calibration time of 80 
seconds \cite{BergSoerensen2004}.

The reason for the discrepancies between all 3 methods is due to what 
is actually being measured. The \textit{ott} estimates are simply 
looking at the differences in the trapping strength along the Cartesian 
axis'. If calibrated over an long enough time frame you would expect 
that the resulting power spectra would exactly mirror the \textit{ott} predictions. However over a short calibration time a QPD can predict
large discrepancies in the trap strength. There is a clear trade off 
in terms of accuracy and computation time as shorter calibration runs 
are computationally more efficient but prone to errors. This is the case
even if we know the exact positional data.

With this in mind, let us consider a symmetric dimer that is optically
trapped by the same Gaussian beam. Not only does the dimer's equilibrium
position change but it is subjected to rotational motion due to the 
beam's angular momentum. This is reflected

\begin{center}
	\captionof{table}{QPD fitting for symmetric dimer}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
		\hline
		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
		& \parbox{1.25cm}{\centering 274} & 285 \\
		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{42.8 \%} &
		\multicolumn{2}{|c|}{12.7 \%} & 
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
Considering the simplicity of a scatterer such as a dimer, one would 
assume that the dynamics of such an object would be relatively easy 
to predict. Simulations of dimers in the Mie regime show that not 
only do they have multiple positions and orientations in which they 
can be trapped but also that their interaction with circularly 
polarised light is heavily dependent on the axial position and 
trapping orientation.

Dimer's have the potential to be used as tunable micro-rotors, being 
simple to synthesise and can be made out of any material of choice.
The rotation demonstrated by dimer's is not accurately described in
previous literature which raises questions on the interactions between
circularly polarised light and spherical aggregates. Being made up of
homogenous materials their is no change in polarisation as expected
by birefringence. Shape birefringence is also not applicable as the 
preferred orientation is perpendicular to the polarisation axis, 
meaning there can be no asymmetry in the dimer's susceptibility.

### 2024-10-25 Latex_files/Thesis/main_win.tex
updates 04-10-2024
updates

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
%\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}
\usepackage{graphicx}
\graphicspath{{C:/Users/xbb16146//dmaciver97/Latex_files/Thesis/figs}}
\usepackage{sidecap}
% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

@@

\usepackage[
backend=biber,
bibstyle=numeric,
citestyle=numeric,
minbibnames=3, maxbibnames=3,
]{biblatex}
\addbibresource{thesis_bib.bib}

@@

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{2}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}
\pretitle{
  \begin{flushright}
  \vspace{-9.5cm}
%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
  \includegraphics[width=5cm]{logo} \\[6cm]
  \end{flushright}
  \begin{center}
  \LARGE
  \posttitle{\end{center}}
  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\chapter{Stuff That Didn't Fit Anywhere Else}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography

>>>>

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
%\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}
\usepackage{graphicx}
\graphicspath{{C:/Users/xbb16146/dmaciver97/Latex_files/Thesis/figs}}
\usepackage{sidecap}
% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

@@

\usepackage[
backend=biber,
bibstyle=ieee,
citestyle=ieee,
minbibnames=3, maxbibnames=3,
]{biblatex}
\addbibresource{thesis_bib.bib}

@@

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}
\pretitle{
  \begin{flushright}
  \vspace{-9.5cm}
%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
  \includegraphics[width=5cm]{logo.pdf} \\[6cm]
  \end{flushright}
  \begin{center}
  \LARGE
  \posttitle{\end{center}}
  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\chapter{Appendix}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimation of fluid shear rates via galvano mirror}
An alternative approach to generating fluid shear is to use a 
galvano-mirror to rapidly move a trapped particle in a bulk 
fluid. While typically galvano and gimble mirrors are used to 
trap multiple particles simultaneously, a single micro sphere 
can be moved quickly through a fluid along a preset path. The 
only limitation on the particle's speed being the ratio of the 
trap stiffness to the drag force. For low Reynolds numbers we 
can assume that the fluid flow is inversely proportional to the
distance from the particle surface (see Eq.~\eqref{eq:birefringent_shear})

For a simple circular path one can estimate the sphere's speed 
by the radius of its path $R$ and the frequency of its orbit 
$\omega$ so that $U = R\omega$. For a more complex path, such 
as an elliptical orbit the curve needs to be parametrised. One 
can describe the position of a sphere on an elliptical path as 
such:
\begin{align}
	r(p) = \left[acos(2\pi p), bsin(2\pi p), 0 \right]
\end{align}

Where a and b are the characteristic radii of an ellipse. If we 
say that $p$ is some dimensionless parameter that describes the 
distance travelled along a trajectory (so $p=0$ is the start of 
the circular orbit and $p=1$ is just as we return to the start).
Then we can rewrite p as $p=t\omega$ where $\omega$ is the period
of orbit. Taking the partial derivative of the position gives 
us the sphere's velocity vector along that path. 
\begin{align}
	v(t) = \frac{\delta r(t)}{\delta t} = \left[-2\pi a\omega \ sin(2\pi t\omega),
	\ 2\pi b\omega \ cos(2\pi t\omega), 0 \right]
\end{align}

In order to compute the particle's speed $U$ we simply take the 
magnitude of our velocity. For low velocities the fluid flow at 
the sphere's surface can be computed based on its velocity.
\begin{align}
	u_r(r)=-|v(t)|\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

Where $r$ is the radial distance to that point. Again taking the partial 
derivative we can get the shear rate for a particle moving through the fluid:
\begin{align}
	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
	\label{eq:galvo_shear}
\end{align}

Moving a silica bead along an circular path can generate significant fluid 
flow around a larger volume compared to comparable micro-rotors. Using 
\eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
equates to moving the silica bead around a circular path with a frequency of 
$100\ Hz$.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.85\linewidth]{galvano_shear_rate.png}
	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
		in bulk water moving at different speeds. The shear rate is calculated 
		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
		in a circular path and so the speed is constant through out its path.}
	\label{fig:galvano_shear}
\end{figure}

From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
mirror a better option for generating high shear rates but also over a larger 
volume. The only limitation to this method is the effective drag experienced 
by the trapped particle. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography
